---
title: "FMIndex_Vignettes"
author: 
  - name: "Cattafi Denis"
    email: "denis.cattafi@mail.polimi.it"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FMIndex_Vignettes}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: sentence
---

## FM-Index {#FM}

This index consists in three data structures that can be used not only for an 
efficient storage of long DNA sequences, but also for count, search or 
extraction of specific patterns.
This functions are not currently implemented, but will be in the future.
The FM-Index can be used for any type of string, but in this package the usage 
is limited to DNA sequences.

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

### Usage Example

```{r setup}
library(FMIndex)
```

```{r,results="hide"}
#Specify the FASTA file, the parameters and the output directory
file <- system.file("extdata", "prova.fasta", package = "FMIndex")
step_tally <- 10
k <- 10
dir <- "./Prova_Results"

# Create the FM-Index in a specified directory
FM_index(file,10,10,dir)
```
### **Output of FM_Index function**

After the execution, 3 CSV files are created in the output directory:

-   `BWT.csv`: Contains the BWT (\ref(BWT)) of the function.
-   `Sparse_Tally.csv`: Contains the sparse tally matrix (\ref(TM)) of the BWT.
-   `Subset_Suffix_Array.csv`: Contains a subset of the Suffix Array (\ref(SA)).

A message of success is returned. 
In case of error, the function stops and an error messages returned

This files can be used for further analysis like pattern count or search, 
that will be implemented in this package in the future


### Clarification
\ref(FM) is the only function to run to obtain the FMIndex as explained in the 
output section.
For completeness, also the subfunctions used to build the index are exported and
accessible.
Their usage is explained in the following vignettes.
\n
\n

#### Subset Suffix Array {#SA}

A data structure of the \ref(FM).
A suffix array is composed of all the suffixes of a string in a 
lexicographical order.
The first data structure of the FM-Index consists in a compressed Suffix Array 
of the original sequence.
The compressed version is simply obtained by selecting the suffixes positions 
that are divisible by a certain value given in input.
The Suffix Array, in this package, is also needed to build the \ref(BWT)

```{r}
# Creates the Suffix Array of a string
DNA_string <- "ACCTGGAC"
SA <- create_SA(DNA_string)
# To select a subset of the suffix array
# NOTE: this is already implemented in the FM_Index function
k <- 2 # set a value in line with the sequence length
SSA <- SA[SA %% k == 0] # Extract 

print(SA)
print(SSA)
```

#### Burrows-Wheeler Transform {#BWT}

A data structure of the \ref(FM).
The Barrows-Wheeler Transform (BWT) is a string of the same length of the 
original sequence, and is saved along with a compressed version of the 
first column of the Burrows-Wheeler Matrix (BWM).
This structures is essential to build the \ref(TM) and to implement the 
pattern recognition funtionality.

```{r}
# Creates the BWT of a DNA_string
DNA_string <- "ACCTGGAC"
bwt <- createBWT(DNA_string)
print(bwt)
```

#### Sparse Tally matrix {#TM}

A data structure of the \ref(FM).
The tally matrix is a cumulative character count of the \ref(BWT).
A sparse version of the tally matrix is useful for efficiently storing this 
structure for long DNA sequences, and is obtained by saving the rows at a 
specific interval defined in input.

```{r}
DNA_string <- "ACCTGGAC"

#creates the sparse tally matrix of the BWT of the DNA string
step <- 2
sparse_tally <- sparse_tally(DNA_string,step)
print(sparse_tally)
```

```{r sessionInfo}
sessionInfo()
```
